---
title: "Simulations of GC content using jump and other models"
output: html_notebook
---

### Background
Here, I will simulate GC content according to a Levy jumps model. The input parameters were manually set equal to the best-fit parameter values obtained by fitting a Levy jumps model to real GC content data using the program levolution.

The way in which the jump rate is encoded in the two programs (levolution and geiger) is different. Levolution gives us jumps per unit time, whereas geiger requires number of jumps. I set the total number of jumps as jump rate*sum of all branch lengths.

I will simulate the data using a modified ex.jumpSimulator function from geiger. This outputs the simulated trait data for every taxa in the given phylogeny and the position (branches) of simulated jumps.

The program also plots a histogram of simulated trait data.

```{r common_stuff }
library(geiger)

data_folder <- "../Input_data/phylogenies/"
sim_res_folder <- "../Results/geigerSim_results/"

GTDB_14k_metadata <- read.delim(file =paste0(data_folder, "GTDB_bac_metadata_r80_pruned0.01.tsv"), header = T, row.names = 1, stringsAsFactors = F)

source("ex_jump_simulator.R")
n_sims <- 1:5
```

#### Rhizobiales
```{r sim_geiger_jumpBM_model_GC_Rhizobiales }
Rhizobiale_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Rhizobiale_GTDB_14k_tree.newick"))

##Here I use the best fit parameters from levolution 
levolution_fit_root <- 64
levolution_fit_sigsq <- 17
levolution_fit_alpha <- 1
levolution_fit_lambda <- 3.5
sum_branch_lengths <- sum(Rhizobiale_GTDB_14k_tree$edge.length)

par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Rhizobiale_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Rhizobiale_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Rhizobiale_GTDB_14k_GC_config1_alpha1_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Rhizobiale_GTDB_14k_GC_config1_alpha1_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Rhodobacterales
```{r sim_geiger_jumpBM_model_GC_Rhodobacterales }
Rhodobacterale_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Rhodobacterale_GTDB_14k_tree.newick"))

levolution_fit_root <- 63
levolution_fit_sigsq <- 36
levolution_fit_alpha <- 1
levolution_fit_lambda <- 2.5
sum_branch_lengths <- sum(Rhodobacterale_GTDB_14k_tree$edge.length)

par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Rhodobacterale_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Rhodobacterale_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Rhodobacterale_GTDB_14k_GC_config1_alpha1_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Rhodobacterale_GTDB_14k_GC_config1_alpha1_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Alpha1
```{r sim_geiger_jumpBM_model_GC_Alpha1s }
Alpha1_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Alpha1_GTDB_14k_tree.newick"))

levolution_fit_root <- 63
levolution_fit_sigsq <- 56
levolution_fit_alpha <- 0.5
levolution_fit_lambda <- 4
sum_branch_lengths <- sum(Alpha1_GTDB_14k_tree$edge.length)

##simulations with higher jump rate
par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Alpha1_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Alpha1_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Alpha1_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Alpha1_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Sphingomonadales
```{r sim_geiger_jumpBM_model_GC_Sphingomonadale_new }
Sphingomonadale_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Sphingomonadale_GTDB_14k_tree.newick"))

levolution_fit_root <- 55
levolution_fit_sigsq <- 21
levolution_fit_alpha <- 1
levolution_fit_lambda <- 2.4
sum_branch_lengths <- sum(Sphingomonadale_GTDB_14k_tree$edge.length)

##simulations with higher jump rate
par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Sphingomonadale_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Sphingomonadale_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Sphingomonadale_GTDB_14k_GC_config1_alpha1_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Sphingomonadale_GTDB_14k_GC_config1_alpha1_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Cytophagales
```{r sim_geiger_jumpBM_model_GC_Cytophagale }
Cytophagale_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Cytophagale_GTDB_14k_tree.newick"))

levolution_fit_root <- 40
levolution_fit_sigsq <- 52
levolution_fit_alpha <- 0.25
levolution_fit_lambda <- 5.6
sum_branch_lengths <- sum(Cytophagale_GTDB_14k_tree$edge.length)

##simulations with higher jump rate
par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Cytophagale_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Cytophagale_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Cytophagale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Cytophagale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Bacteroidale
```{r sim_geiger_jumpBM_model_GC_Bacteroidale }
Bacteroidale_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Bacteroidale_GTDB_14k_tree.newick"))

levolution_fit_root <- 40
levolution_fit_sigsq <- 71
levolution_fit_alpha <- 0.25
levolution_fit_lambda <- 2.4
sum_branch_lengths <- sum(Bacteroidale_GTDB_14k_tree$edge.length)

##simulations with higher jump rate
par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Bacteroidale_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Bacteroidale_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Bacteroidale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Bacteroidale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Flavobacteriale
```{r sim_geiger_jumpBM_model_GC_Flavobacteriale }
Flavobacteriale_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Flavobacteriale_GTDB_14k_tree.newick"))

levolution_fit_root <- 40
levolution_fit_sigsq <- 15
levolution_fit_alpha <- 1
levolution_fit_lambda <- 6.5
sum_branch_lengths <- sum(Flavobacteriale_GTDB_14k_tree$edge.length)

##simulations with higher jump rate
par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Flavobacteriale_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Flavobacteriale_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Flavobacteriale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Flavobacteriale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Pseudomonadale
```{r sim_geiger_jumpBM_model_GC_Pseudomonadale }
Pseudomonadale_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Pseudomonadale_GTDB_14k_tree.newick"))

levolution_fit_root <- 52
levolution_fit_sigsq <- 56
levolution_fit_alpha <- 0.25
levolution_fit_lambda <- 5.18
sum_branch_lengths <- sum(Pseudomonadale_GTDB_14k_tree$edge.length)

##simulations with higher jump rate
par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Pseudomonadale_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Pseudomonadale_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Pseudomonadale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Pseudomonadale_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Enterobacterales
```{r sim_geiger_jumpBM_model_GC_Enterobacterales }
Enterobacterales_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Enterobacterales_GTDB_14k_tree.newick"))

levolution_fit_root <- 46
levolution_fit_sigsq <- 56
levolution_fit_alpha <- 0.5
levolution_fit_lambda <- 3.85
sum_branch_lengths <- sum(Enterobacterales_GTDB_14k_tree$edge.length)

##simulations with higher jump rate
par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Enterobacterales_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Enterobacterales_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Enterobacterales_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Enterobacterales_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```

#### Betaproteo
```{r sim_geiger_jumpBM_model_GC_Betaproteo }
Betaproteo_GTDB_14k_tree <- read.tree(file = paste0(data_folder, "Betaproteo_GTDB_14k_tree.newick"))

levolution_fit_root <- 58
levolution_fit_sigsq <- 37
levolution_fit_alpha <- 0.5
levolution_fit_lambda <- 6
sum_branch_lengths <- sum(Betaproteo_GTDB_14k_tree$edge.length)

par(mfrow=c(2,3))
hist(GTDB_14k_metadata[Betaproteo_GTDB_14k_tree$tip.label,"gc_percentage"], main="actual", xlab="GC")
for(sim_n in n_sims){
  sim_jBM <- ex.jumpsimulator(phy = Betaproteo_GTDB_14k_tree, alpha = levolution_fit_root, sigmasq.brown = levolution_fit_sigsq, sigmasq.jump = levolution_fit_alpha*levolution_fit_sigsq, jumps = levolution_fit_lambda*sum_branch_lengths)
  
  hist(sim_jBM$hist$phenotype, main=paste0("sim",sim_n), xlab="GC")
  write.table(x = sim_jBM$dat, file = paste0(sim_res_folder, "Betaproteo_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".traits"), quote = F, row.names = T, col.names = F, sep = "\t")
  write.csv(x = sim_jBM$hist, file = paste0(sim_res_folder, "Betaproteo_GTDB_14k_GC_config1_alpha",levolution_fit_alpha,"_geigerSim", sim_n, ".details"), quote = F, row.names = F)
}
```