---
title: "precision recall curves for simulated jumps"
output: html_notebook
---

### Background
Here, I will analyze the inference (by levolution) of jumps simulated according to a Levy jumps model.

```{r preliminaries }
library(phytools)
library(rncl)

main_dir <- "../Results/geigerSim_results/"
res_dir <- main_dir
out_dir <- res_dir
```

### Summary of precision-recall
Here, I will regenerate precision-recall curves for all clades analysed upto now-
```{r summary_precision_recall }
clades_oi <- c("Sphingomonadale", "Rhizobiale", "Rhodobacterale", "Alpha1", "Bacteroidale", "Cytophagale", "Flavobacteriale", "Pseudomonadale", "Enterobacterales", "Betaproteo")
p_folders <- c(rep("Alphaproteo",4), rep("Bacteroidetes",3), rep("Gammaproteo",3))
alphas_oi <- c(1,1,1,0.5,0.25,0.25,1,0.25,0.5,0.5,2,1,0.1)
config_oi <- 1
n_sims <- 1:5

pp_range <- c(seq(0,0.7,0.1),seq(0.75,0.9,0.05),seq(0.91,1,0.01))
PR_df <- as.data.frame(matrix(data=NA, nrow = length(clades_oi)*length(pp_range), ncol=4, dimnames = list(NULL, c("clade", "pp", "precision", "recall"))))
  
cnt <- 1
for(clade_n in 1:length(clades_oi)){
  precision_oi <- recall_oi <- c()
  clade_oi <- clades_oi[clade_n]
  parent_oi <- p_folders[clade_n]
  alpha_oi <- alphas_oi[clade_n]
  
  for(pp_cutoff in pp_range){
    comparison_matrix <- matrix(data = 0, nrow = length(n_sims), ncol = 4, dimnames = list(n_sims,c("TP", "TN", "FP", "FN")))
    
    for(sim_n in n_sims){
      ##posterior probabilities inferred by levolution on simulated trait data
      tree_pp <- read.newick(file = paste0(res_dir,clade_oi,"_GTDB_14k_GC_config", config_oi, "_alpha", alpha_oi,"_geigerSim", sim_n, ".post"))
      
      ##actual jump locations in simulated data
      actual_jump_details <- read.csv(paste0(res_dir, clade_oi, "_GTDB_14k_GC_config", config_oi, "_alpha", alpha_oi,"_geigerSim", sim_n, ".details"), header = T, stringsAsFactors = F)
      actual_jump_edges <- which(actual_jump_details$jumps>0)
      actual_jump_col <- sapply(actual_jump_details$effect_Jump[actual_jump_edges], function(x) ifelse(x<0,"red","blue"))
      actual_jump_size <- sapply(actual_jump_details$effect_Jump[actual_jump_edges], function(x) sqrt(abs(x)))
      actual_jump_size <- actual_jump_size/max(actual_jump_size)
      
      comparison_matrix[as.character(sim_n),"TP"] <- length(intersect(which(tree_pp$edge.length>=pp_cutoff), actual_jump_edges)) ##True positives
      comparison_matrix[as.character(sim_n),"TN"] <- length(intersect(which(tree_pp$edge.length<pp_cutoff), setdiff(1:length(tree_pp$edge.length), actual_jump_edges))) ##True negatives
      comparison_matrix[as.character(sim_n),"FP"] <- length(intersect(which(tree_pp$edge.length>=pp_cutoff), setdiff(1:length(tree_pp$edge.length), actual_jump_edges))) ##False positives
      comparison_matrix[as.character(sim_n),"FN"] <- length(intersect(which(tree_pp$edge.length<pp_cutoff), actual_jump_edges)) ##False negatives
    } ##loop over sims

    inference_summary <- colSums(comparison_matrix)
    precision_oi <- inference_summary["TP"]/(inference_summary["TP"]+inference_summary["FP"])
    recall_oi <- inference_summary["TP"]/(inference_summary["TP"]+inference_summary["FN"])
    
    PR_df[cnt, ] <- c(clade_oi, pp_cutoff, precision_oi, recall_oi)
    cnt <- cnt + 1
  } ##loop over pps
} #loop over clades

write.table(PR_df, file = paste0(res_dir, "levolution_precision_recall_data.tab"), quote = F, row.names = F)
```

#### Plotting
Lets plot the precision-recall data for all clades
```{r summary_plots_precision-recall }
library(ggplot2)
library(ggrepel)
require(cowplot)
library(tidyverse)

PR_df <- read.table(file = paste0(res_dir, "levolution_precision_recall_data.tab"), stringsAsFactors = F, header = T)
PR_df$clade <- factor(x = PR_df$clade, levels = unique(PR_df$clade))

PR_df$threshold <- "normal"
PR_df$threshold_text <- "normal_text"
pp_thresholds <- c(0.75,0.75,0.75,0.95,0.95,0.95,0.75,0.9,0.9,0.9)
for(clade_n in 1:length(clades_oi)){
  clade_oi <- clades_oi[clade_n]
  pp_threshold <- pp_thresholds[clade_n]
  PR_df$threshold[PR_df$clade==clade_oi & PR_df$pp==pp_threshold] <- "highlight"
  PR_df$threshold_text[PR_df$clade==clade_oi & PR_df$pp==pp_threshold] <- "highlight_text"
}

p <- ggplot(subset(PR_df, clade %in% clades_oi[1:10] & pp %in% c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.75, 0.9, 0.95, 1)), aes(recall*100, precision*100, color=pp)) + geom_point()
p <- p + geom_point(data = subset(PR_df, clade %in% clades_oi[1:10] & threshold=="highlight"), aes(recall*100, precision*100, label=pp), color="orange")
p <- p + scale_x_continuous(name = "recall (%)", limits = c(0,100)) + scale_y_continuous(name = "precision (%)", limits = c(0,100))
p <- p + geom_text_repel(data = subset(PR_df, clade %in% clades_oi[1:10] & threshold=="highlight"), inherit.aes = F, aes(x=recall*100,y=precision*100,label=pp), color="orange", size=4.5, show.legend = F)
p <- p + scale_color_continuous()
p <- p + facet_wrap(~ clade, ncol = 3, scales = "fixed")
p <- p + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(colour = "black"), legend.position = c(1, 0), legend.justification = c(1, 0), legend.key.width = unit(x = 1, units = "line"), legend.title.align = 0, axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20), strip.text.x = element_text(size=14))
ggsave(filename = paste0(res_dir, "PRC_all_facets.svg"), plot = p, dpi = 300, width = 8, height = 11, units = "in")
ggsave(filename = paste0(res_dir, "PRC_all_facets.png"), plot = p, dpi = 300, width = 8, height = 11, units = "in")
```

#### Variation in prcision and recall at chosen thresholds
We may be interested in knowing how much does precision and recall vary for every individual simulation, rather than pooled data. Lets assess how much precision and recall vary at the chosen threshold for every clade.
```{r summary_PRC_variation }
clades_oi <- c("Sphingomonadale", "Rhizobiale", "Rhodobacterale", "Alpha1", "Bacteroidale", "Cytophagale", "Flavobacteriale", "Pseudomonadale", "Enterobacterales", "Betaproteo")
p_folders <- c(rep("Alphaproteo",4), rep("Bacteroidetes",3), rep("Gammaproteo",3))
alphas_oi <- c(1,1,1,0.5,0.25,0.25,1,0.25,0.5,0.5)
pp_thresholds <- c(0.75,0.75,0.75,0.95,0.95,0.95,0.75,0.9,0.9,0.9)
config_oi <- 1
n_sims <- 1:5

library(phytools)

PR_df <- as.data.frame(matrix(data=NA, nrow = length(clades_oi)*5, ncol=6, dimnames = list(NULL, c("clade", "pp_threshold", "sim_n", "num_jumps", "precision", "recall"))))

PR_var_df <- as.data.frame(matrix(data=NA, nrow = length(clades_oi), ncol=10, dimnames = list(NULL, c("subclade", "pp_threshold", "min_num_jumps", "max_num_jumps", "overall_precision", "min_precision", "max_precision", "overall_recall", "min_recall", "max_recall"))))

cnt <- 1
for(clade_n in 1:length(clades_oi)){
  clade_oi <- clades_oi[clade_n]
  parent_oi <- p_folders[clade_n]
  pp_cutoff <- pp_threshold <- pp_thresholds[clade_n]
  alpha_oi <- alphas_oi[clade_n]
  
  comparison_matrix <- matrix(data = 0, nrow = length(n_sims), ncol = 4, dimnames = list(n_sims,c("TP", "TN", "FP", "FN")))
 
  for(sim_n in n_sims){
    tryCatch({
      tree_pp <- read.newick(file = paste0(res_dir,clade_oi,"_GTDB_14k_GC_config", config_oi, "_alpha", alpha_oi,"_geigerSim", sim_n, ".post"))
    }, error=function(e){
      print(e)
    }, finally = {
      actual_jump_details <- read.csv(paste0(res_dir, clade_oi, "_GTDB_14k_GC_config", config_oi, "_alpha", alpha_oi,"_geigerSim", sim_n, ".details"), header = T, stringsAsFactors = F)
      actual_jump_edges <- which(actual_jump_details$jumps>0)
      actual_jump_col <- sapply(actual_jump_details$effect_Jump[actual_jump_edges], function(x) ifelse(x<0,"red","blue"))
      actual_jump_size <- sapply(actual_jump_details$effect_Jump[actual_jump_edges], function(x) sqrt(abs(x)))
      actual_jump_size <- actual_jump_size/max(actual_jump_size)
      
      tp_oi <- length(intersect(which(tree_pp$edge.length>=pp_cutoff), actual_jump_edges)) ##True positives
      tn_oi <- length(intersect(which(tree_pp$edge.length<pp_cutoff), setdiff(1:length(tree_pp$edge.length), actual_jump_edges))) ##True negatives
      fp_oi <- length(intersect(which(tree_pp$edge.length>=pp_cutoff), setdiff(1:length(tree_pp$edge.length), actual_jump_edges))) ##False positives
      fn_oi <- length(intersect(which(tree_pp$edge.length<pp_cutoff), actual_jump_edges)) ##False negatives
      
      comparison_matrix[as.character(sim_n),"TP"] <- tp_oi
      comparison_matrix[as.character(sim_n),"TN"] <- tn_oi
      comparison_matrix[as.character(sim_n),"FP"] <- fp_oi
      comparison_matrix[as.character(sim_n),"FN"] <- fn_oi
      
      precision_oi <- tp_oi/(tp_oi+fp_oi)
      recall_oi <- tp_oi/(tp_oi+fn_oi)
      
      PR_df[cnt, ] <- c(clade_oi, pp_cutoff, as.character(sim_n), length(actual_jump_edges), round(precision_oi, digits = 2), round(recall_oi, digits = 2))
      cnt <- cnt + 1
    })
  } ##loop over sims
  
  inference_summary <- colSums(comparison_matrix)
  precision_pooled <- inference_summary["TP"]/(inference_summary["TP"]+inference_summary["FP"])
  recall_pooled <- inference_summary["TP"]/(inference_summary["TP"]+inference_summary["FN"])
    
  
  PR_var_df[clade_n,] <- c(clade_oi, pp_cutoff, min(as.integer(PR_df[PR_df$clade==clade_oi, "num_jumps"]), na.rm = T), max(as.integer(PR_df[PR_df$clade==clade_oi, "num_jumps"]), na.rm = T), round(precision_pooled,2), min(as.numeric(PR_df[PR_df$clade==clade_oi, "precision"]), na.rm = T), max(as.numeric(PR_df[PR_df$clade==clade_oi, "precision"]), na.rm = T), round(recall_pooled,2), min(as.numeric(PR_df[PR_df$clade==clade_oi, "recall"]), na.rm = T), max(as.numeric(PR_df[PR_df$clade==clade_oi, "recall"]), na.rm = T))
}

write.table(x = PR_var_df, file = paste0(main_dir, "all_clades_PR_variation.tab"), row.names = PR_var_df$subclade)
```